---
title: "Geocoding Market Coordinates"
author: "Yanbingyi"
date: "2021-04-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.align = "center")
```

## Introduction

This report uses the names and the address of markets that were contained in the food price data from Global Food Prices Database [(WFP)](https://data.humdata.org/dataset/wfp-food-prices) to geocode their coordinates in longitude and latitude. The geocoding service that used in this report is provided by Google Map. To facilitate the future analysis, each market will be matched to its country code (ISO3) and will be assigned to a unique market ID. At the end of the report, a csv file that included the market names, address, ISO3, coordinates, and unique market ID will be created. A point shapefile will also be created for visualization and further analysis.


## Geocode Market Location

```{r}
library(tidyverse)
## Read data
wfp_foodprice <- read.csv("data/wfpvam_foodprices.csv")

## Extract market names and address
market.address <- wfp_foodprice %>% 
  select(adm0_name, adm1_name, mkt_name) %>% 
  mutate(mkt_country = paste(mkt_name, adm0_name, sep = ", "),
         mkt_province = paste(mkt_name, adm1_name, sep = ", ")) %>% 
  distinct(mkt_country, mkt_province, .keep_all = T)
```

```{r}
library(ggmap)
```


```{r include=FALSE}
# Set API key
register_google("AIzaSyDOnGSpn3jCOvFBxA-4cywo4oD-NdEpseU")
```


```{r}
## Set API key
## register_google("API")
## Geocoding by market name+country
Coord <- market.address %>% 
  mutate_geocode(mkt_country)

## Some markets can't be located by market name+country 
## Try market name+province
Coord.missing <- Coord %>% 
  filter(is.na(lon)) %>% 
  select(-lon, -lat) %>% 
  mutate_geocode(mkt_province) 

## Combine two dataframe
Address <- Coord %>% 
  left_join(Coord.missing, by = c("adm0_name", "adm1_name", "mkt_name", "mkt_country", "mkt_province")) %>% 
  mutate(longitude = case_when(is.na(lon.x) ~ lon.y, T ~ lon.x),
         latitude = case_when(is.na(lat.x) ~ lat.y, T ~ lat.x)) %>% 
  select(-lon.x, -lon.y, -lat.x, -lat.y)

## Markets can't be located
Missing <- Address %>% filter(is.na(longitude))
nrow(Missing)
head(Missing)
```

Since we can't locate these markets in map, these markets will not be included in future analysis


## Match with ISO3 and assign market ID

```{r}
ISO3 <- read.csv("data/COUNTRY CODE.csv")
colnames(ISO3)[1] <- "Country"
colnames(Address)[1] <- "Country"

Address <- Address %>% 
  left_join(ISO3, by = "Country")

## Manually enter the ISO3 for markets can't be matched
## this is because some countries have different names 
ISO3.missing <- Address %>% 
  filter(is.na(`Alpha.3.code`)) %>% 
  select(Country, `Alpha.3.code`) %>% 
  distinct()
ISO3.missing
```


The current officially assigned ISO 3166-1 alpha-3 codes and their corresponding country names can be found [here](https://unstats.un.org/unsd/tradekb/Knowledgebase/50347/Country-Code)


```{r}
ISO3.missing$Alpha.3.code <- c("IND", "CIV", "COD", "IRN", "MDA", "SSD", "PSE", "TZA")

Address <- Address %>% 
  filter(!is.na(longitude)) %>% 
  mutate(ISO3 = as.character(Alpha.3.code)) %>% 
  select(-Alpha.3.code) %>% 
  left_join(ISO3.missing, by = "Country") 

Address$ISO3[is.na(Address$ISO3)] <- Address$Alpha.3.code[is.na(Address$ISO3)]

## Final result
Mkt_info <- Address %>% 
  select(ISO3, Country, Province = adm1_name, Market = mkt_name, Address = mkt_country, latitude, longitude)

```

There are a few previously geocoded markets. We will combine the old list to our new list

```{r}
old.list <- read.csv("data/market_coord.csv")
head(old.list)

old.list <- old.list %>% 
  left_join(ISO3, by = c("ISO" = "Alpha.3.code"))

## Check if there is any issue with the country name
old.list[is.na(old.list$Country),]

## "SSD" is South Sudan
##  We will manually fill in the country name
old.list$Country <- as.character(old.list$Country)
old.list$Country[old.list$ISO == "SSD"] <- "South Sudan"

Coord.old <- old.list %>% 
  select(ISO3 = ISO, Country, Market = MARKET) %>% 
  mutate(Address = paste(Market, Country, sep = ", ")) %>% 
  mutate_geocode(Address)

## Check if there is any geocoding issue
Coord.old[is.na(Coord.old$lat), ]

## Will remove this observation
Coord.old <- Coord.old %>% 
  filter(!is.na(lon)) %>% 
  mutate(Province = NA) %>% 
  select(ISO3, Country, Province, Market, Address, latitude = lat, longitude = lon)
```


```{r}
## Combine two dataframe
## Some markets in the old list are overlapped with the new list
## We will only keep one observation if duplicated
remove.dup <- data.frame(lat = Coord.old$latitude, lon = Coord.old$longitude, lat.check = NA, lon.check = NA)
remove.dup$lat.check <- remove.dup$lat %in% Mkt_info$latitude
remove.dup$lon.check <- remove.dup$lon %in% Mkt_info$longitude

Mkt_info <- Mkt_info %>% 
  bind_rows(Coord.old[(remove.dup$lat.check + remove.dup$lon.check) == 0,]) %>% 
  arrange(Country, Province, Market) %>% 
  mutate(ID = paste0(ISO3, seq(length(Country))))

```


```{r}
write.csv(Mkt_info, file = "output/Market_Address.csv")
```

